---
title: "Troubleshooting FAQ"
description: >
  Learn how to contribute to our project.
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Troubleshooting FAQ}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

*Purpose: This document compliments Sweave2Rmd [Contribute](https://bioconductor.github.io/sweave2rmd/articles/contribute.html) 
page. It is a FAQ inventory for common issues you may run into during the 
document conversion process. For more context on the project, see the 
[Sweave2Rmd project page](https://bioconductor.github.io/sweave2rmd/index.html).* 

## Common Questions: 

### *Question:* Can I Add a Footnote in the YAML title or author fields? 

*Answer:* Ordinarily, footnotes are put inside the square brackets after a caret.
e.g., `^[This is a footnote.]`. However, due to a bug, footnotes in the YAML 
title or author fields no longer run. If attempted, you will receive a a LaTeX 
error. If you are converting a vignette and notice that the author byline 
includes a footnote, as does the Bioconductor vignette [Random Numbers in BiocParallel](https://docplayer.net/228745333-Random-numbers-in-biocparallel.html), 
there is a work around. In the YAML, add three drop down lines for the author. 
Include all or a combination of the following: name, affiliation, and email. See
below. 

```{r eval=FALSE}
author:
  name: Martin Morgan
  affiliation: Roswell Park Cancer Institute, Buffalo, NY
  email: https://www.roswellpark.org/martin-morgan
``` 

### *Question:* How can I handle duplicate chunks?
Background info: When working with R markdown files, duplicate chunks tend to prevent knitting and make the `Biocheck` fail.

*Answer*: Look for the exact duplicste files and ask the reviewer if its okay to make some changes. For example if the chunk name is 'duplicate' you can make the other chunk name `duplicate1` or `duplicate_` as long as it does not change the meaning or make it hard for someone to understand the code. If you do not understand the code or document well, always ask the maintainer or reviewers for what to do.

```{r eval=FALSE}
author:
  name: Beryl Kanali
  affiliation: xxx
``` 
